Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from the biclaR project

### 2023-11-28 paper/PaperCEUS/PaperCEUS.Rmd
start PT car legs fig
Remove xxx

---
title: "Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from the XXX project"
# subtitle: "Reproducible methods, results and actionable evidence from the XXX project"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_XXX_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 10% of trips could be made by bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the PT leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

@@

To achieve the cycling targets set by the Portuguese national cycling strategy for 2025 and 2030 (4% and 10%, respectively) [@ENMAC], the Lisbon's Metropolitan Department of Transport introduced *XXX*[^1], a decision support tool that facilitates the planning, design, and development of a metropolitan cycling network (ref X).
<!-- [@felix2023]. -->

@@

*XXX* builds on the Propensity to Cycle Tool[^2] (PCT), a web application and research project funded by the UK's Department for Transport in 2015 which launched nationally in 2017 as part of the government's Cycling and Walking Investment Strategy.
The PCT initially used only origin-destination data for commuting trips as the basis of estimates of cycling potential at zone, route and route network levels [@lovelace2017].
The PCT has been extended to include cycling potential for travel to school in England [@goodman2019] and other trip types in other countries.[^3]
However, to the best of our knowledge, this is the first time that the method has been integrated with public transport data using multi-modal routing to estimate the potential and benefits of multi-stage cycling and PT trips.

@@

The conversion of avoided emissions into avoided welfare loss and respective monetary valuation was based on the EU Guide to Cost-benefit Analysis [@EuropeanCommission2014] and the best up-to-date reference values for the various gases [@EuropeanCommission2014; @bickel2006; @UNITE]: 8.44 €/ton for CO, 2,867.85 €/ton for NO~X~, 340,969.27 €/ton for PM10, 7,169.62 €/ton for VOC and 35.85 €/ton for CO~2~eq.
The social impacts are in avoided premature mortality. This result is finally monetized using the *Statistical Value of Life* for Portugal: €3,055,358/fatality [@ANSR2021].
We updated all the monetary reference values of the literature based on the annual inflation rate in Portugal for 2022[^7], and our 10-years estimations assumed a discount rate of 5% and inflation of 3%.

@@

The information available at *XXX* tool -- an open access website -- can be downloaded and used with any GIS software. This allows practitioners to, for example, gain insights into which potential cycling connections have the highest socio-environmental impacts, quantified in tons of avoided CO~2~eq emissions, or in long term social benefits. <!-- move to introduction? -->

@@

## Research data {.unnumbered}
The data and the code to produce XXX and this paper are available at: https://github.com/[*blind*].

>>>>

---
title: "Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from the biclaR project"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 10% of trips could be made by bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the PT leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

@@

To achieve the cycling targets set by the Portuguese national cycling strategy for 2025 and 2030 (4% and 10%, respectively) [@ENMAC], the Lisbon's Metropolitan Department of Transport commissioned *biclaR*[^1], a decision support tool that facilitates the planning, design, and development of a metropolitan cycling network (ref X).
<!-- [@felix2023]. -->

@@

*biclaR* builds on the Propensity to Cycle Tool[^2] (PCT), a web application and research project funded by the UK's Department for Transport in 2015 which launched nationally in 2017 as part of the government's Cycling and Walking Investment Strategy.
The PCT initially used only origin-destination data for commuting trips as the basis of estimates of cycling potential at zone, route and route network levels [@lovelace2017].
The PCT has been extended to include cycling potential for travel to school in England [@goodman2019] and other trip types in other countries.[^3]
However, to the best of our knowledge, this is the first time that the method has been integrated with public transport data using multi-modal routing to estimate the potential and benefits of multi-stage cycling and PT trips.

@@

The conversion of avoided emissions into avoided welfare loss and respective monetary valuation was based on the EU Guide to Cost-benefit Analysis [@EuropeanCommission2014] and the best up-to-date reference values for the various gases [@EuropeanCommission2014; @bickel2006; @UNITE]: 8.44 €/ton for CO, 2,867.85 €/ton for NO~X~, 340,969.27 €/ton for PM10, 7,169.62 €/ton for VOC and 35.85 €/ton for CO~2~eq.
The social impacts are in avoided premature mortality. This result is finally monetized using the *Statistical Value of Life* for Portugal: €3,055,358/fatality [@ANSR2021].
We updated all the monetary reference values of the literature based on the annual inflation rate in Portugal for 2022[^7], and our 10-years estimations assumed a discount rate of 5% and inflation of 3%. 
See [Research Data](#research-data) for all the input values we used.

@@

The results can can be downloaded from the *biclaR* and used with any GIS software. This allows practitioners to, for example, gain insights into which potential cycling connections have the highest socio-environmental impacts, quantified in tons of avoided CO~2~eq emissions, or in long term social benefits. <!-- move to introduction? -->

@@

## Research data {.unnumbered}
The data and the code to reproduce the results will be made available upon publication.
